Low-Cost Data Deduplication for Virtual Machine Backup in Cloud Storage

### 2013-05-14 hotstorage13/arch.tex
re

\item {\bf Separation of duplicate detection and data backup.}
%Request accumulation and partition-based deduplication.}
The second level detection requires a global comparison of fingerprints.
% of stored chunks with the new chunks scanned during the backup process. 
%Bloom filters and caching allows some of content 
%fingerprints to be stored on the cheap disks~\cite{bottleneck08}. Such an optimization is good 
%for inline deduplication where chunks scanned need to be determined to be duplicate or not instantly without 
%waiting while  it memory consumption is still significant in competing with stand virtual 
%machine activities.  
Our approach is to perform duplicate detection first before actual data backup.
That requires a prescanning of  data  blocks in dirty VM segments, which
%  and after that, real backup  for those non-duplicates is conducted. 
does incur an extra  round of VM reading.
% while avoiding inline deduplication and leading to a much smaller resource requirement.  
During VM prescanning, detection requests are accumulated.
Aggregated duplicate requests can be processed partition by partition. 
%We also accumulate the delete requests and perform them partition by partitions. 
Since each partition corresponds to a small portion of global content index, 
memory cost to process detection requests within a partition is small.
%the cost to process detection requests within a partition is significantly smaller than processing all 
%the requests globally.

\item {\bf Buffered data redistribution in parallel duplicate detection}.  
Let {\em global index} be the meta data containing the fingerprint values of unique blocks of snapshots
in all VMs and  the reference pointers to the location of raw data.
%A duplicate detection request for a content block needs to be compared with the global index to determine if this
%block is non-duplicate. If it is a duplicate, return the corresponding reference pointer.
%We conduct parallel processing of detection requests in all machines involved. 
A logical way to distribute detection requests among machines is to partition based on 
fingerprint values of content blocks.
% in a  disjointed manner.
Initial data  blocks follows the VM distribution among machines and the detected duplicate summary 
should be collected following the same distribution.
Therefore, there are two all-to-all data redistribution operations involved.
One is to map detection requests from VM-based distribution to fingerprint based distribution.  
Another one  is to map duplicate summary from fingerprint-based distribution to VM based distribution.  
The redistributed data needs to be accumulated on the disk to reduce the use of memory.
To minimize the disk seek cost, outgoing or incoming data exchange messages are buffered to 
bundle small messages.
Given there are $p\times q$ partitions where $p$ is the number of machines and $q$ is the number of fingerprint-based partitions
at each machine, space per each buffer  is small under the memory constraint for large $p$ or $q$ values.
This counteracts the effort of seek cost reduction.  
%Then there will be a large number of storage  IO operations involved, suffering the huge seek overhead.
We have designed an efficient data exchange and disk data buffering  scheme to address this.

@@

%\item 
Stage 3 is to perform real backup one VM at a time.
The system loads the duplicate summary of a VM, 
reads  dirty segments of a VM, and outputs non-duplicate blocks to the final backup 
storage. Additionally, the global index on each machine is updated with the meta data of new chunk blocks. 
When a segment is not dirty, the system only needs to output the segment meta data such as a reference pointer. 
There is an option to directly read dirty blocks instead of fetching a dirty segment which can include duplicate
blocks. Our experiment shows that it is faster to read dirty segments in the tested workload.
Another issue is that during global index update after new block creation in the backup storage,
the system  may find some  blocks with the same fingerprints have been 
created redundantly. For example, two different VM blocks that have the same  fingerprint are not detected
because  the  global index has not contained such a fingerprint yet. 
The redundancy is discovered and logged during the index update and can be repaired
periodically when necessary.  Our experience is that there is a redundancy during the initial snapshot backup and once 
that is repaired, the percentage of redundant blocks due to concurrent processing  is insignificant.

>>>>

\item {\bf Separation of duplicate detection and data backup.}
%Request accumulation and partition-based deduplication.}
The second level detection requires a global comparison of fingerprints.
% of stored chunks with the new chunks scanned during the backup process. 
%Bloom filters and caching allows some of content 
%fingerprints to be stored on the cheap disks~\cite{bottleneck08}. Such an optimization is good 
%for inline deduplication where chunks scanned need to be determined to be duplicate or not instantly without 
%waiting while  it memory consumption is still significant in competing with stand virtual 
%machine activities.  
Our approach is to perform duplicate detection first before actual data backup.
That requires a prescanning of  dirty VM segments, which
%  and after that, real backup  for those non-duplicates is conducted. 
does incur an extra  round of VM reading.
% while avoiding inline deduplication and leading to a much smaller resource requirement.  
During VM prescanning, detection requests are accumulated.
Aggregated duplicate requests can be processed partition by partition. 
%We also accumulate the delete requests and perform them partition by partitions. 
Since each partition corresponds to a small portion of global content index, 
memory cost to process detection requests within a partition is small.
%the cost to process detection requests within a partition is significantly smaller than processing all 
%the requests globally.

\item {\bf Buffered data redistribution in parallel duplicate detection}.  
Let {\em global index} be the meta data containing the fingerprint values of unique snapshot blocks 
in all VMs and  the reference pointers to the location of raw data.
%A duplicate detection request for a content block needs to be compared with the global index to determine if this
%block is non-duplicate. If it is a duplicate, return the corresponding reference pointer.
%We conduct parallel processing of detection requests in all machines involved. 
A logical way to distribute detection requests among machines is based on 
fingerprint values of content blocks.
% in a  disjointed manner.
Initial data  blocks follows the VM distribution 
among machines 
and the detected duplicate summary 
should be collected following the same distribution.
Therefore, there are two all-to-all data redistribution operations involved.
One is to map detection requests from VM-based distribution to fingerprint based distribution.  
Another one  is to map duplicate summary from fingerprint-based distribution to VM based distribution.  
The redistributed data needs to be accumulated on the disk to reduce the use of memory.
To minimize the disk seek cost, outgoing or incoming data exchange messages are buffered to 
bundle small messages.
Given there are $p\times q$ partitions where $p$ is the number of machines and $q$ is the number of fingerprint-based partitions
at each machine, space per each buffer  is small under the memory constraint for large $p$ or $q$ values.
This counteracts the effort of seek cost reduction.  
%Then there will be a large number of storage  IO operations involved, suffering the huge seek overhead.
We have designed an efficient data exchange and disk data buffering  scheme to address this.

@@

%\item 
Stage 3 is to perform real backup.
% one VM at a time.
The system loads the duplicate summary of a VM, 
reads  dirty segments of a VM, and outputs non-duplicate blocks to the final backup 
storage. Additionally, the global index on each machine is updated with the meta data of new chunk blocks. 
When a segment is not dirty, the system only needs to output the segment meta data such as a reference pointer. 
There is an option to directly read dirty blocks instead of fetching a dirty segment which can include duplicate
blocks. Our experiment shows that it is faster to read dirty segments in the tested workload.
Another issue is that during global index update after new block creation in the backup storage,
the system  may find some  blocks with the same fingerprints have been 
created redundantly. For example, two different VM blocks that have the same  fingerprint are not detected
because  the  global index has not contained such a fingerprint yet. 
The redundancy is discovered and logged during the index update and can be repaired
periodically when necessary.  Our experience is that there is a redundancy during the initial snapshot backup and once 
that is repaired, the percentage of redundant blocks due to concurrent processing  is insignificant.

### 2013-05-14 hotstorage13/background.tex
re

%The previous work for storage backup has extensively used  data deduplication techniques can eliminate redundancy globally among different files from different users.
Backup systems have been developed to use content fingerprints to identify duplicate
content~\cite{venti02,Rhea2008}.
Offline deduplication is used in ~\cite{EMC,NetAppOffline} to remove previously written duplicate blocks during idle time.
%,NGmiddleware2011}.
%Today's commercial data backup systems (e.g. from EMC and NetApp)
%\cite{emc_avamar}\cite{datadomain_whitepaper}
%use a variable-size chunking algorithm to detect duplicates in file data~\cite{similar94,hydrastor09}.
Several techniques have been proposed to speedup searching of duplicate
fingerprints. For example, the data domain method ~\cite{bottleneck08} 
uses  an in-memory Bloom filter and a prefetching cache for data blocks  which may be
accessed.  An improvement to this work with parallelization is in ~\cite{MAD210,DEBAR}.
%Still such methods use a significant amount of memory resource for filtering and caching.
Such methods use dedicated nodes for deduplication, which add cost to cloud infrastructure and is less scalable.
%NG et al.~\cite{ NGmiddleware2011}  use
%a related filtering technique for integrating deduplication in Linux  file system and the memory
%consumed is up to 2GB for a single machine. That is still too big in our context discussed below.
%Lillibridge et al.~\cite{sparseindex09} break list of chunks
%into large segments, the chunk IDs in each incoming segment are sampled and the segment is
%deduplicated by comparing with the chunk IDs of only a few carefully selected backed up segments.
%These are segments that share many chunk IDs with the incoming segment with high probability.
The approximation techniques are studied in~\cite{extreme_binning09,Guo2011,WeiZhangIEEE}  
%Deepavali et al.~\cite{extreme_binning09} and Zhang et al.~\cite{WeiZhangIEEE}  
to reduce memory requirement with a tradeoff of the reduced deduplication ratio.
%use fingerprint-based file similarity  and group similar files into the same physical location (bins) to deduplicate against each other.
%That leads  to a smaller amount of memory usage for storing meta data in fingerprint
%lookup  
In comparison, this paper focuses on  full deduplication without approximation.

>>>>

%The previous work for storage backup has extensively used  data deduplication techniques can eliminate redundancy globally among different files from different users.
Backup systems have been developed to use content fingerprints to identify duplicate
content~\cite{venti02,Rhea2008}.
Offline deduplication is 
used in ~\cite{EMC,NetAppOffline} to remove previously written duplicate blocks during idle time.
%,NGmiddleware2011}.
%Today's commercial data backup systems (e.g. from EMC and NetApp)
%\cite{emc_avamar}\cite{datadomain_whitepaper}
%use a variable-size chunking algorithm to detect duplicates in file data~\cite{similar94,hydrastor09}.
Several techniques have been proposed to speedup searching of duplicate
fingerprints. For example, the data domain method ~\cite{bottleneck08} 
uses  an in-memory Bloom filter and a prefetching cache for data blocks  which may be
accessed.  An improvement to this work with parallelization is in ~\cite{MAD210,DEBAR}.
Still such methods use  a significant amount of memory resource for filtering and caching.
As discussed in Section~\ref{sect:intro}, 
there is no dedicated resource for deduplication in our targeted setting and low memory usage 
is required so that the resource impact to other cloud services is minimized.

%NG et al.~\cite{ NGmiddleware2011}  use
%a related filtering technique for integrating deduplication in Linux  file system and the memory
%consumed is up to 2GB for a single machine. That is still too big in our context discussed below.
%Lillibridge et al.~\cite{sparseindex09} break list of chunks
%into large segments, the chunk IDs in each incoming segment are sampled and the segment is
%deduplicated by comparing with the chunk IDs of only a few carefully selected backed up segments.
%These are segments that share many chunk IDs with the incoming segment with high probability.
The approximation techniques are studied in~\cite{extreme_binning09,Guo2011,WeiZhangIEEE}  
%Deepavali et al.~\cite{extreme_binning09} and Zhang et al.~\cite{WeiZhangIEEE}  
to reduce memory requirement with a tradeoff of the reduced deduplication ratio.
%use fingerprint-based file similarity  and group similar files into the same physical location (bins) to deduplicate against each other.
%That leads  to a smaller amount of memory usage for storing meta data in fingerprint
%lookup  
In comparison, this paper focuses on  full deduplication without approximation.
%We also take advantages of the fact that in a VM cloud environment,
%the virtual device driver can easily keep track if  large data
%segments have been modified using dirty bits and such information can avoid sending
%unmodified data segments for deduplication, which significantly saves cost.
